### 2025-01-31 scripts/Paper-analysis.Rmd
chi-squared test for top depleted motifs

>>>>

### All motifs testing for over-representation

Go through values of k for across PTU averages.

```{r}
# k= 4
# 6-bp palindromes
palindromes = read.csv('~/Downloads/trieste/k4.txt', header=F)$V1
results.df = read.csv('../test_ptus_k4_scores.csv',header=T)
# Add PTUs as column names
colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
# Normalise scores by the number of plasmids within PTU?
n.plasmids = table(lr.no.duplicates.subset$PTU)
results.df.norm = results.df %>%
mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
# Plot all palindromes
results.df.long = results.df.norm %>%
pivot_longer(cols = starts_with("PTU"),
names_to = "score_type",
values_to = "score_avg")
results.df.norm.average = results.df.long %>% group_by(kmer) %>%
summarise(median=median(score_avg)) %>%
arrange(median)
# Take the top 1% depleted
results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.05))),] # not enough for 0.01
results.df.norm.average.top.depleted$palindrome = ifelse(results.df.norm.average.top.depleted$kmer %in% palindromes,
"palindrome",
"non-palindrome")
# Read in all RM targets
all_rm_targets = read.csv('../data/all_rebase_targets_250131_6bp_less_deambiguated.txt', header=F, sep=' ')
all_rm_targets_4bp =all_rm_targets$V1[all_rm_targets$V2==4]
results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
all_rm_targets_4bp,
"target", "non-target")
table(results.df.norm.average.top.depleted$palindrome)
table(results.df.norm.average.top.depleted$target)
chisq.test(matrix(c(3, 10, 13, 230 ), nrow=2))
chisq.test(matrix(c(4, 9, 99, 144 ), nrow=2))

### And for k=5
results.df = read.csv('../test_ptus_k5_scores.csv',header=T)
# Add PTUs as column names
colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
# Normalise scores by the number of plasmids within PTU?
n.plasmids = table(lr.no.duplicates.subset$PTU)
results.df.norm = results.df %>%
mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
results.df.long = results.df.norm %>%
pivot_longer(cols = starts_with("PTU"),
names_to = "score_type",
values_to = "score_avg")
results.df.norm.average = results.df.long %>% group_by(kmer) %>%
summarise(median=median(score_avg)) %>%
arrange(median)
# Take the top 1% depleted
results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),] # Top 5% since so few
all_rm_targets_5bp =all_rm_targets$V1[all_rm_targets$V2==5]
results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
all_rm_targets_5bp,
"target", "non-target")
table(results.df.norm.average.top.depleted$target)
chisq.test(matrix(c(5, 6, 277, 736 ), nrow=2))

# k =6
# 6-bp palindromes
palindromes = read.csv('~/Downloads/trieste/k6.txt', header=F)$V1
results.df = read.csv('../test_ptus_k6_scores.csv',header=T)
# Add PTUs as column names
colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
# Normalise scores by the number of plasmids within PTU?
n.plasmids = table(lr.no.duplicates.subset$PTU)
results.df.norm = results.df %>%
mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
# Plot all palindromes
results.df.long = results.df.norm %>%
pivot_longer(cols = starts_with("PTU"),
names_to = "score_type",
values_to = "score_avg")
results.df.norm.average = results.df.long %>% group_by(kmer) %>%
summarise(median=median(score_avg)) %>%
arrange(median)
# Take the top 1% depleted
results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),]
results.df.norm.average.top.depleted$palindrome = ifelse(results.df.norm.average.top.depleted$kmer %in% palindromes,
"palindrome",
"non-palindrome")
# Read in all RM targets
all_rm_targets = read.csv('../data/all_rebase_targets_250131_6bp_less_deambiguated.txt', header=F, sep=' ')
all_rm_targets_6bp =all_rm_targets$V1[all_rm_targets$V2==6]
results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
all_rm_targets_6bp,
"target", "non-target")
table(results.df.norm.average.top.depleted$palindrome)
table(results.df.norm.average.top.depleted$target)
chisq.test(matrix(c(7, 34, 865-7, 4096-865-34 ), nrow=2))
chisq.test(matrix(c(4, 37, 60, 4055 ), nrow=2))
